Dynamics of Anthropomorphism in Human-Robot Interaction

### 2014-11-07 paper.tex
Start to shorten the paper, update the pictures

\newcommand{\eg}{{\textit{e.g.~}}}
\newcommand{\etal}{{\textit{et al.~}}}
\newcommand{\ie}{{\textit{i.e.~}}}
\newcommand{\etc}{{\textit{etc.~}}}
\newcommand{\vs}{{\textit{vs.~}}}

@@

\begin{figure}[b]
        \centering
        \begin{subfigure}[t]{0.48\columnwidth}
                \includegraphics[width=\textwidth]{roomba-activities.png}
                \caption{Activities related to Roomba, n = 9 households}
                \label{fig:roomba-activities}
        \end{subfigure}%
        \hspace{0.5cm} 
        \begin{subfigure}[t]{0.48\columnwidth}
                \includegraphics[width=\textwidth]{roomba-perception.pdf}
                \caption{People's perception of Roomba, n = 15 participants}
                \label{fig:roomba-perception}
        \end{subfigure}
    \caption{Usage and perception of the Roomba vacuum-cleaning robot over a 6
    months period. Qualitative data suggest that people's activities related to
    the robot as well as their perception of the robot change over time.}
    \label{fig:roomba}
\end{figure}

@@

Further, we would like to investigate the link between the variances in
anthropomorphism and the correlating cognitive states.

@@

 \subsection{The lexical field of anthropomorphism}

Originally, the term \textbf{anthropomorphism} comes from the Greek
\textit{anthropos} for ``man'' (or ``human'') and \textit{morphe} for ``form /
structure'' (or ``shape''). Anthropomorphism is generally understood as people's
tendency to perceive a non-human agent as if it had some human characteristics
or traits. How a person perceives an agent is crucial, as it further determines
how he/she relates to, and behaves towards the agent, or how he/she may behave in
light of this agent \citep{epley_when_2008}. An anthropomorphic relationship is
often characterized by the fact that the human attributes human-unique
characteristics to the non-human agent. Since anthropomorphism becomes
observable through these attributions, some definitions claim that
anthropomorphism is not the \textit{perception} but rather the
\textit{attribution} of human qualities to non-human agents.  Ascribing
thoughts, emotions, and intentions are given as examples, and as the most
prominently anthropomorphized non-human agents, animals and robots are mentioned
\citep{duffy_anthropomorphism_2003,schmitz_concepts_2011}.

Apart from anthropomorphism, there also exists the term \textbf{animism}, which
can be defined as the attribution of life to the nonliving. The study of animism
was pioneered by Piaget in the 1920s with his work on \textit{The Child's
Conception of the  World}. He found that children tend to regard many inanimate
objects as capable of sensations, emotions and intentions, and called this view
``animism''. Since then, children's inferences about the world have been studied
intensively, and there is an agreement that children's views and concepts of
living, animal, and plant evolve as they grow up. Often synonymously used to
animism is the term \textbf{zoomorphism}, which similarly applies to the cases
when non-lifelike instances / artifacts are associated with animalistic
attributes, excluding human-specific traits.  As ``being alive'' is not a
human-unique property but applies to all living beings, the attribution of life
may not be labeled as anthropomorphism (in a narrow understanding of it).
Another possibility to illustrate how anthropomorphism differs from attributions
of animacy is to look at paraphrases of anthropomorphism. That is to say,
anthropomorphism can alternatively be understood as \textbf{humanization} or
\textbf{personification} of anything other than a human being.\footnote{There
exist also the phenomenons inverse to anthropomorphism, which can be found
in the literature as \textbf{dehumanization}
\citep{haslam_dehumanization:_2006}, or \textbf{mechanomorphism}
\citep{caporael_anthropomorphism_1986}.} Robots can
both be animized and anthropomorphized, but often the difference is not made.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.75\columnwidth]{in_our_heads.pdf}
    \caption{[ALTERNATIVE TO FIGURE \ref{fig:anthropofig}] Anthropomorphism is a
        subjective, psychological phenomenon which arises in a social context, during an
        interaction between a human agent and a non-human agent. One can take
        two main perspectives when trying to explain anthropomorphism: a
        human-centered and an artifact-centered perspective.
        Several external factors influence the anthropomorphism phenomenon.
        Person- and artifact-related characteristics, but also related to the
        context of use as well as the actual time spent interacting with the
        robot (interaction history).}

    \label{fig:anthropofig-alt}
\end{figure}

@@

Most experiments, and even more when they investigate long-term interactions,
are carried out with robots exhibiting relatively simple behaviors. In our
model, the maximum level of anthropomorphism \AntMax is reached with the novelty
effect, and then decreases. A natural question is whether we can design robots
such that \AntMax remains high, possibly higher than $\ant[t_{novelty~peak}]$
(note that to know if sustaining strong anthropomorphic interaction is desirable
is entirely another question, see Section \ref{sec:ethics}).

Intuitively, the behavioral complexity of the robot, and in particular, the
ability for the robot to evolve over time and \emph{adapt its behavior to the
user} as the user familiarizes his/herself with the robot (\ie,
\emph{co-familiarize}), seems to play a key role.

@@

\subsection{General observations}

By adopting both a cognitive and a dynamic perspective on the anthropomorphic
bonds that establish between a robot and a human during an interaction, some
observations can be raised.

For instance, we propose that the human \emph{adapts} iteratively to the robot
by restating its cognitive model of the robot. Could our two models be
conversely relied on to have the robot itself iteratively adapting its behavior?
In particular, we hypothesize that the so-called novelty effect represents a
peak in anthropomorphic manifestations, followed by mostly deception-driven
refinements of the cognitive behavioral model of the robot.  Could we imagine
that the robot pro-actively \emph{verbalizes} its own limits, in a timely manner
(likely before the end of the novelty effect). And how would this affect the
cognitive models built by the human?

One related question we plan to elaborate on is the effects of \emph{mutual
explicitation of the mental model of the other agent}: a human and a robot
interact on a given task, and after a while, we interrupt the task and ask each
of the agents to explicit the mental model it has built of its partner
(emotional state, beliefs, intentions, etc.). Depending on the cognitive stages
the two agents are in, we may expect varying accuracy levels in the produced
mental models.

Another question raised by these models relates to their transposition to
human-human interaction: while concepts like \emph{novelty effect} are not
directly meaningful when analyzing human-human interaction, insights from
cognitive sciences (and in particular social psychology) regarding how we, as
humans, cope with unexpected behaviors from peers, or on the dynamics of mental
model refinements, could bring interesting perspectives with possible
applications to better manage long-term human-robot interactions.

@@

\subsection{Ethical aspects of anthropomorphism and human-like form of robots}
\label{sec:ethics}

After all discussion, there remains the question how far a sustained
anthropomorphic (human social) interaction between a human and a non-human can
be favored, regarding ethical aspects. We need to rethink carefully the
consequences that emotional bonds between humans and robots can have. They might
be beneficial on one hand but can also hold negative implications. How far
should a robot's design encourage us to lastingly project feelings or intentions
to it? Why should a machine even react to us as if it had emotions? Can this be
the reality we want to live in? What happens as soon as the artificial agent is
considered a \textit{moral agent} \citep{sullins_when_2006}? Also, the use of
\textit{socially interactive robots} in contexts involving elderly care or in
rehabilitation and with autistic children is a controversially discussed topic
\cite{robins_robots_2005}. Do we want robots to comfort us; robots that make us
eventually even falling in love with them?  The anthropomorphic design of robots
is critical. Why do we even need this? The HRI / robotics community needs to
discuss these and other questions.

>>>>

\usepackage{xspace}
\newcommand{\eg}{\textit{e.g.}\xspace}
\newcommand{\etal}{\textit{et al.}\xspace}
\newcommand{\ie}{\textit{i.e.}\xspace}
\newcommand{\etc}{\textit{etc.}\xspace}
\newcommand{\vs}{\textit{vs.}\xspace}

@@

\begin{figure}[b]
        \centering
        \begin{subfigure}[t]{0.48\columnwidth}
                \includegraphics[height=7cm]{roomba-activities}
                \caption{Activities related to Roomba, \textit{n = 9 households}}
                \label{fig:roomba-activities}
        \end{subfigure}%
        \hspace{0.5cm} 
        \begin{subfigure}[t]{0.48\columnwidth}
                \includegraphics[height=7cm]{roomba-ratings}
                \caption{People's perception of Roomba, \textit{n = 15
                participants}}
                \label{fig:roomba-perception}
        \end{subfigure}
    \caption{Usage and perception of the Roomba vacuum-cleaning robot over a 6
    months period. Qualitative data suggest that people's activities related to
    the robot as well as their perception of the robot change over time.}
    \label{fig:roomba}
\end{figure}

@@

@@

@@

%
%\begin{figure}
%    \centering
%    \includegraphics[width=0.75\columnwidth]{in_our_heads.pdf}
%    \caption{[ALTERNATIVE TO FIGURE \ref{fig:anthropofig}] Anthropomorphism is a
%        subjective, psychological phenomenon which arises in a social context, during an
%        interaction between a human agent and a non-human agent. One can take
%        two main perspectives when trying to explain anthropomorphism: a
%        human-centered and an artifact-centered perspective.
%        Several external factors influence the anthropomorphism phenomenon.
%        Person- and artifact-related characteristics, but also related to the
%        context of use as well as the actual time spent interacting with the
%        robot (interaction history).}
%
%
%    \label{fig:anthropofig-alt}
%\end{figure}
%

@@

Most experiments, and even more when they investigate long-term interactions,
are carried out with robots exhibiting relatively simple behaviors. In our
model, the maximum level of anthropomorphism \AntMax is reached with the novelty
effect, and then decreases. A natural question is whether we can design robots
such that \AntMax remains high, possibly higher than $\ant[t_{novelty~peak}]$
(note that to know if sustaining strong anthropomorphic interaction is desirable
is entirely another question).

Intuitively, the behavioral complexity of the robot, and in particular, the
ability for the robot to evolve over time and \emph{adapt its behavior to the
user} as the user familiarizes his/herself with the robot (\ie
\emph{co-familiarize}), seems to play a key role.

@@

@@
